---
title: "Survival Group Project"
author: "Yujin Zhang"
date: '2022-11-08'
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(survival)
library(ggplot2)
library(ggfortify)

knitr::opts_chunk$set(echo = TRUE)
```

# Load Data

```{r, message=FALSE}
lung
data(cancer, package="survival")

cancer_df=cancer
head(cancer_df, 7)
```

Dataset Attributes Overview:  
*inst:	    Institute Code  
*time:	    Survival time in days  
*status:	  Censoring status; 1= censored; 2 = dead  
*age:	    Age of the subject in years  
*ph.ecog:  Ecog performance score; 0: good, 5: dead  
*ph.karno: Karnofsky performance score; 0: bad; 100: good rated by physician  
*pat.karno:Karnofsky performance score as rated by patient  
*meal.cal:	Calories consumed at meals  
*wt.loss:	Weight loss in last six months  

Next, clean the data
```{r}
map(cancer_df, ~sum(is.na(.))) # count missing value

# add event variable where censor=0 and death=1
cancer_df = cancer_df %>% 
  mutate( event = ifelse(status > 1, 1, 0),
          sex = as.factor(sex)
  )
```


# Kaplan-Meier survival estimation & Nelson-Aalen hazard estimation 

## Kaplan-Meier survival estimation
```{r}
km_fit = survfit(Surv(time, event) ~ 1, data = cancer_df)
summary(km_fit)

#plot M-K 
autoplot(km_fit)

# observe median survival time
print(km_fit) 
```
The median survival time of this KM estimator is 310 days.

## look at survival curves by sex.
```{r}
km_sex = survfit(Surv(time, event) ~ sex, data = cancer_df)
summary(km_fit)

#plot M-K 
autoplot(km_sex)    #plot(km_sex, xlab="Days", main = 'Kaplan Meyer Plot')

# observe median survival time
print(km_sex)


```
The median survival time for male(sex=1) is 270 days, and the median survival time for female(sex=2) is 426 days.



## Nelson-Aalen hazard estimation 
```{r}
nelsonaalen <- function(data, timevar, statusvar) {
  if (!is.data.frame(data)) {
    stop("Data must be a data frame")
  }
  timevar <- as.character(substitute(timevar))
  statusvar <- as.character(substitute(statusvar))
  time <- data[, timevar, drop = TRUE]
  status <- data[, statusvar, drop = TRUE]

  hazard <- survival::basehaz(survival::coxph(survival::Surv(time, status) ~ 1))
  idx <- match(time, hazard[, "time"])
  hazard[idx, "hazard"]
}

hazard = nelsonaalen(cancer_df, time, event)
plot(x = cancer_df$time, y = hazard, ylab = "Cumulative hazard", xlab = "Time")
```

# Log-rank hypothesis test
First, apply log-rank test on sex


# Cox Proportional Hazards Model
